\documentclass[acmsmall,screen,nonacm]{acmart}
\citestyle{acmauthoryear}
\settopmatter{printfolios=true}
\emergencystretch=1em

\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}

% ------------------------------------------------------------------
% Kimiya code style (shared with the Kimiya paper)
% ------------------------------------------------------------------
\definecolor{kwcolor}{RGB}{30,60,130}     % core Kimiya keywords: blue
\definecolor{extcolor}{RGB}{140,30,90}    % OUR extension keywords: magenta
\definecolor{cmtcolor}{RGB}{110,110,110}
\definecolor{typecolor}{RGB}{120,60,20}
\lstdefinelanguage{Kimiya}{
  % class 1: the ORIGINAL Kimiya keyword set, verbatim from the Kimiya paper
  morekeywords={retry,until,budget,panel,paraphrase_prompts,under,gen,judge,
    select,check,commit,abstain,if,then,else,case,repeat,forall,variant,
    context,domain,preserve,allow_loss,decreases,not,otherwise,when},
  % class 2: the world-effecting EXTENSION this paper proposes (S6) --
  % typeset distinctly so a reader can see at a glance which language a
  % construct belongs to
  morekeywords=[2]{act,observe,settle,within,compensate},
  morecomment=[l]{--},
  morecomment=[l]{\#},
  morestring=[b]",
  sensitive=true
}
\lstset{
  language=Kimiya,
  basicstyle=\small\ttfamily,
  keywordstyle=\color{kwcolor}\bfseries,
  keywordstyle=[2]\color{extcolor}\bfseries,
  commentstyle=\color{cmtcolor}\itshape,
  stringstyle=\color{typecolor},
  columns=fullflexible, keepspaces=true, showstringspaces=false,
  xleftmargin=1.2em, aboveskip=0.6em, belowskip=0.6em,
  escapeinside={(*}{*)}
}

% ------------------------------------------------------------------
% Kimiya macros (subset used here)
% ------------------------------------------------------------------
\newcommand{\semeq}{\mathrel{\sim}}
\newcommand{\semeqk}[1]{\mathrel{\sim_{#1}}}
\newcommand{\entails}{\mathrel{\models}}
\newcommand{\entailsk}[1]{\mathrel{\models_{#1}}}
\newcommand{\den}[1]{\llbracket #1 \rrbracket}
\newcommand{\Dist}{\mathcal{D}_{\le 1}}
\newcommand{\gen}{\mathsf{gen}}
\newcommand{\checkk}{\mathsf{check}}
\newcommand{\judge}{\mathsf{judge}}
\newcommand{\retry}{\mathsf{retry}}
\newcommand{\select}{\mathsf{select}}
\newcommand{\commitc}{\mathsf{commit}}
\newcommand{\abstain}{\mathsf{abstain}}
\newcommand{\actc}{\mathsf{act}}
\newcommand{\observe}{\mathsf{observe}}
\newcommand{\settle}{\mathsf{settle}}
\newcommand{\Skip}{\mathsf{skip}}
\newcommand{\ctx}{\kappa}
\newcommand{\loss}{\ell}
\newcommand{\conf}{\theta}
\newcommand{\budget}{B}
\newcommand{\Oracle}{\mathcal{O}}
\newcommand{\true}{\mathsf{tt}}
\newcommand{\false}{\mathsf{ff}}
\newcommand{\Prob}{\Pr}
\newcommand{\bolt}{\maltese}
\newcommand{\ir}[3]{\dfrac{\;#2\;}{\;#3\;}\ \textsc{(#1)}}
\newcommand{\worldstore}{\mathcal{W}}
\newcommand{\ext}[1]{\textcolor{extcolor}{#1}}

\AtEndPreamble{%
  \theoremstyle{acmdefinition}%
  \newtheorem{principle}[theorem]{Design Principle}%
  \theoremstyle{acmplain}%
  \newtheorem{claimx}[theorem]{Claim}%
  \newtheorem{remark}[theorem]{Remark}%
}

\begin{document}

\title{GUI Testing as a Kimiya Program: Deterministic Scripts, Calibrated
Perception, and a Path to Verified Interaction with GUI, Network, and Cloud
Software}

\author{Mahmudul Faisal Al Ameen}
\affiliation{\institution{(draft manuscript)}\city{Tokyo}\country{Japan}}

\begin{abstract}
We report a concrete instance of a general thesis: an end-to-end test of
interactive software---graphical, networked, cloud-backed---is naturally
expressed as a \emph{Kimiya program}, a computation whose control flow is
deterministic and whose meaning-laden steps are calibrated language-model
instruments. We built a testing harness in which mouse, keyboard, vision,
closed-loop audio, network fault injection, and production-cloud
verification operate together over a desktop application and two
concurrent browsers, and applied it to a deployed product, finding ten
production bugs. We then observe that the harness had, unknowingly,
implemented the Kimiya discipline: its element-locator is a $\select$
instrument, its screen-assertions are $\judge$ guards under implicit
purposes, its on-screen value reads are grounded $\gen$, and its
database, cloud-API, image-hash, and audio-waveform oracles are exactly
the trusted $\checkk$ kernel that keeps judge error from becoming a false
certificate. We make the correspondence precise, transcribe a
representative scenario into Kimiya surface syntax, and---using an
operating point \emph{measured} from the harness's own run
artifacts ($\beta \ge 0.976$, $\alpha \le 0.16$, per-call locate recall
$\ge 0.975$)---derive a numeric reliability certificate for a passing
test. The correspondence is not perfect: GUI testing stresses exactly the
extension Kimiya defers, computation that \emph{mutates an external
world under adversarial timing}, and we propose the constructs a
world-effecting Kimiya needs ($\actc$/$\observe$/$\settle$, world-frame
retry, timed cost) together with the proof obligations they carry.
Finally, we recast the artifact as an executable guideline: a playbook by
which an autonomous language-model agent can test arbitrary GUI software
under the same discipline. The larger aim is a route to \emph{formal
verification of interactive systems} that neither pretends the world is
deterministic nor lets the model's judgment go uncalibrated---using
exactly as much determinism, and exactly as much language model, as each
step requires, with the split made auditable by a Kimiya certificate.
\end{abstract}

\keywords{GUI testing, LLM-as-judge, program logic, calibrated oracles,
end-to-end testing, verified interaction, Kimiya}

\maketitle

% ==================================================================
\section{Introduction}
\label{sec:intro}
% ==================================================================

End-to-end testing of a graphical application is the only test layer that
exercises the product as a user does, and the one teams most often
abandon. Selector-bound scripts decay with every interface change;
cross-platform products need a separate harness per surface; and whole
classes of behavior---audio, multi-user synchronization, degraded
networks, the live production backend---are declared out of scope because
no convenient handle exists. Two recent capabilities reopen the problem.
Vision-language models locate interface elements from natural-language
descriptions reliably enough for production use, and ``computer-use''
agents can operate arbitrary interfaces through screenshots and synthetic
input. The dominant response is to build \emph{autonomous testing
agents} that explore an application and improvise actions. Autonomy,
however, is the wrong property for a regression test: an improvising
agent does not run the same test twice, cannot be reviewed as an
artifact, and its failures resist attribution.

This paper argues, and demonstrates on a real system, a different
position. An interactive test wants to be \emph{deterministic in its
control flow and calibrated in its perception}. The steps that decide
\emph{what to do}---open this view, present these pages, then end the
talk---must be a fixed, reviewable script. The steps that require
\emph{meaning}---find the button described as ``the red End
button,'' judge whether the screen shows a live-presentation state, read
the join code off a dialog---are irreducibly semantic and belong to a
language model. And crucially, the \emph{verdict} must never rest on
perception alone: a screen that merely looks right must not be able to
pass a test. This is precisely the shape of a \emph{Kimiya}
program~\citep{kimiya}: deterministic structure around a triad of
datasheet-carrying instruments---$\gen$ (produce), $\judge$
(discriminate), $\select$ (find)---with a trusted kernel $\checkk$ for
what is decidable, and a program logic that transfers reliability across
judged guards at \emph{measured} error rates.

We did not set out to instantiate Kimiya. We built a testing harness for
SeenSlide---a deployed slide-sharing product with a PyQt5 desktop
capture application, a FastAPI/PostgreSQL cloud backend, and a Svelte
web viewer with real-time collaboration---because the product needed
testing that its unit suite could not provide. The harness drives, with
one step vocabulary: a native desktop application including a GTK file
dialog; two \emph{different} real browsers concurrently under distinct
user accounts; a synthetic fullscreen presenter standing in for a human;
an operating-system-level virtual microphone whose audio is verified,
after crossing a WebRTC mesh, at the speaker output; a fault-injecting
network relay; and a live production cloud. On eighteen scenarios it
found ten production bugs, three of them invisible on screen and caught
only by non-visual oracles, including a database-pool defect that
returned HTTP 500 to \emph{every authenticated user} and a three-bug
cascade in which each fix exposed the next.

Only afterward did the correspondence become clear: the harness had
implemented the Kimiya discipline without naming it. That correspondence
is this paper's subject.

\paragraph{Contributions.}
\begin{itemize}
\item \textbf{The correspondence (\S\ref{sec:harness}--\S\ref{sec:as-kimiya}).}
A precise mapping from an interactive testing harness to Kimiya: locator
$=\select$, screen-assertion $=\judge$ under an implicit purpose, value-
read $=$ grounded $\gen$, and the database/cloud/image-hash/audio oracles
$=$ the trusted $\checkk$ kernel. We transcribe a representative two-user
scenario into Kimiya surface syntax.
\item \textbf{A measured datasheet (\S\ref{sec:datasheet}).} We calibrate
the vision locator from the harness's own run artifacts---58 runs, 604
model calls---obtaining a conservative operating point, and use it to
derive a \emph{numeric} reliability certificate for a passing test, in
the style of the Kimiya logic. This turns a green run from a boolean into
a priced guarantee, and separates the claims a test certifies at
kernel-certainty from those it certifies only at the judge's error rate.
\item \textbf{An extension for effectful, timed interaction
(\S\ref{sec:extend}).} GUI testing stresses exactly what Kimiya's core
defers: computation over an external, mutating, adversarially timed
world. We propose the constructs a world-effecting Kimiya needs---
$\actc$/$\observe$/$\settle$, a world-frame retry rule, and timed
cost---and state the proof obligations each introduces.
\item \textbf{An executable playbook (\S\ref{sec:playbook}).} We recast
the artifact as a guideline an autonomous language-model agent can follow
to test arbitrary GUI software under this discipline, so the method
generalizes beyond the one system we tested.
\item \textbf{A verification agenda (\S\ref{sec:verification}).} We argue
that this decomposition is a viable route to formal verification of
interactive, networked, cloud-backed software: use exactly as much
determinism and as much language model as each step requires, and make
the division auditable by a certificate whose meaning depends only on
declarations and datasheets.
\end{itemize}

We are explicit about status: the guarantees are conditional on the
oracle model (judges being calibratable instruments), the datasheet is
measured on one system's interface families, and the extension of
\S\ref{sec:extend} is a design, not yet a mechanized metatheory. What is
solid is the empirical core---the harness, the ten bugs, the measured
operating point---and the demonstration that a real interactive test
\emph{is} a Kimiya program.

% ==================================================================
\section{Kimiya in Brief}
\label{sec:background}
% ==================================================================

We recall only what this paper uses; \citet{kimiya} give the full
calculus, denotational semantics, logic, and metatheory.

Kimiya treats a language model as a processor whose basic operations act
on meaning. Its instruction core is three \emph{oracles}, each carrying a
\emph{datasheet}---an empirically measured operating point:
\begin{itemize}
\item $x := \gen\langle S\rangle(p)$ produces an $S$-schema-valid value
from prompt $p$; it is \emph{untrusted} (validity $\nu_S$ is syntactic;
no semantic property of the output is assumed).
\item $\judge\langle k,\tau\rangle\,\psi$ evaluates a semantic predicate
$\psi$ by $k$ oracle samples at acceptance threshold $\tau$, with a
measured confusion matrix: false-accept rate $\alpha=\Prob[\true\mid\psi\
\text{false}]$, true-accept $\beta=\Prob[\true\mid\psi\ \text{true}]$,
and intra-panel correlation $\rho$.
\item $x := \select\langle\rho_r\rangle(q,w)$ retrieves from store $w$
with datasheet-stamped recall $\rho_r=\Prob[i\in x\mid i\ \text{relevant}]$.
\item $\checkk\,\varphi$ is the \emph{trusted kernel}: a decidable
predicate (schema validation, a test, a proof check) at certainty $1$,
cost $0$.
\end{itemize}
Control flow is internalized in judged conditionals ($\mathbf{if}\
\judge\ \psi\ \mathbf{then}\ C_1\ \mathbf{else}\ C_2$) and \emph{verified
retry} ($\retry\ C\ \mathbf{until}\ J\ \mathbf{budget}\ B$: rerun $C$ from
the loop's entry state until judgment $J$ passes or the budget is spent,
whereupon the program \emph{abstains} to an explicit outcome $\bolt$).
Value identity is a \emph{purpose-indexed judged equivalence}
$a\semeqk{\ctx}b$: reflexive, symmetric, and deliberately
\emph{non-transitive}---a \emph{tolerance relation} of which classical
equality is the zero-tolerance special case. A purpose
$\ctx=\langle\mathit{domain},\mathit{preserve},\mathit{allow\_loss}\rangle$
declares what an identification must preserve and what it may lose;
``the same'' is always ``the same for a purpose.''

Programs denote maps in a cost-graded subdistribution monad---divergence
is honest missing mass, budget exhaustion an explicit observable
$\bolt$, so \emph{failure is never silent}. On top sits a distributional
Hoare logic with triples $\{P\}\,C\,\{Q\}_{\ge\conf}@\budget$: from
precondition $P$, program $C$ commits in a $Q$-state with probability at
least $\conf$ within budget $\budget$. The rules are parameterized by
datasheet quantities; the two we use directly are:
\[
\ir{Judge-If}{
(\alpha,\beta)\ \text{from }d\quad
\{P\wedge\psi\}C_1\{Q\}_{\ge\conf_1}\quad
\{P\wedge\neg\psi\}C_2\{Q\}_{\ge\conf_2}
}{
\{P\}\ \mathbf{if}\ \judge\langle k,\tau\rangle\psi\ \mathbf{then}\ C_1\
\mathbf{else}\ C_2\ \{Q\}_{\ge\min(\beta\conf_1,(1-\alpha)\conf_2)}
}
\]
\[
\ir{Select}{
(\rho_r,\pi)\ \text{from }d\quad \forall i.\ \{P\wedge i\in x\}C\{Q(i)\}_{\ge\conf}
}{
\forall i\in\mathrm{Rel}(q,w).\ \{P\}\ x:=\select\langle\rho_r\rangle(q,w);C\
\{Q(i)\}_{\ge\rho_r\conf}
}
\]
(For readability we omit Kimiya's cost terms from both conclusions, and
write $(\alpha,\beta)$ for what the full rules consume as conservative
\emph{effective} ends $(\alpha_{\mathrm{eff}},\beta_{\mathrm{eff}})$ of a
deployed panel configuration; our judges are single-sample, where the two
coincide.) Two structural facts matter below. First, a judged guard transfers
reliability across an \emph{imperfect observation of a latent
truth}---the branch genuinely executes under $\psi$ or $\neg\psi$, but all
a run obtains is the judge's record. Second, a claim proved through
$\checkk$ carries certainty $1$; a claim proved through $\judge$ carries
only $\beta$ (or $1-\alpha$). Kimiya keeps these two grades distinct, and
that distinction is the hinge of our analysis.

% ==================================================================
\section{The Harness}
\label{sec:harness}
% ==================================================================

We summarize the artifact; the correspondence to Kimiya is \S\ref{sec:as-kimiya}.

\paragraph{Split of concerns.} Scenarios are prewritten step lists (a YAML
DSL). A \emph{runner} interprets them; a \emph{locator} wraps a
mid-tier vision model (Claude Opus~4.8, invoked headlessly) behind three
primitives---\textsf{locate} (natural-language description
$\to$ bounding box and click point), \textsf{judge} (``does the screen
show $X$?'' $\to$ boolean with a one-sentence reason), and \textsf{read}
(``what value is displayed here?'' $\to$ a string). Locate results are
cached by (screen size, description); a replay mode reruns whole
scenarios from the cache with zero model calls. The vision model appears
in exactly these three primitives and nowhere else; it never chooses the
next step.

\paragraph{The dimensions.} Beyond click and type, the harness supplies
what interactive testing usually excludes. A \emph{presenter} renders PDF
pages fullscreen on a chosen monitor under a schedule, standing in for
the human who presents. A \emph{virtual microphone} (a PipeWire loopback
whose playback side is a device-class source) becomes the system default
input; text-to-speech (piper) speaks one line per slide, and on the
output side the harness records the default sink's monitor and asserts
waveform peaks---including an inverted mode that proves \emph{silence}. In
the WebRTC voice test the speech feeds only the loopback microphone, so
any energy at the speaker sink can only have crossed the mesh from the
sending browser. A \emph{relay} routes the application's cloud traffic
through a local HTTP proxy; killing it is a clean network outage. Two
\emph{browser actors} bind names to real, user-prepared Edge and Firefox
windows by title; steps carrying a target crop the screenshot to that
window and offset clicks into it, so two viewers of the same page cannot
be confused.

\paragraph{Oracles.} Every scenario ends in checks that are not visual:
structural queries against the application's SQLite; perceptual hashing of
each stored slide against the source PDF pages (best-match, with an order
constraint, because deduplication legitimately merges near-identical
pages); the \emph{public} cloud endpoint---what the audience's viewer
actually renders---for slide and viewer counts; audio-waveform peaks; and
regex evidence in the application's own logs. Artifacts of every run---a
film-strip HTML report with the located box and click point on each
screenshot, the application log, a consistent copy of its database, and
recorded audio---are retained.

\paragraph{Yield.} Across eighteen scenarios the harness found ten
production bugs, all fixed and shipped, summarized in
Table~\ref{tab:bugs}. Every one is an \emph{integration} bug beyond the
reach of the project's 65-test unit suite; three were invisible on screen
and caught purely by oracles; and bugs 8--10 form a cascade in which each
fix exposed the next.

\begin{table}[t]\small
\caption{The ten production bugs found by the harness. ``Oracle'' = caught
by a non-visual check; the rest surfaced through scripted interaction plus
a judged or read step.}
\label{tab:bugs}
\begin{tabular}{@{}rll@{}}
\toprule
\# & Defect (one line) & Caught by \\
\midrule
1 & config key never read $\Rightarrow$ all slide data in volatile \texttt{/tmp} & DB oracle \\
2 & slide-marker wiring only in the cloud branch of voice recording & audio + log oracle \\
3 & hardcoded white input under a dark-theme white text token & screen judge \\
4 & slide gate armed too late; exempted the app's own frames & cloud oracle + report \\
5 & session slide counter incremented, never decremented on delete & cloud oracle \\
6 & session row never persisted; Sessions view showed ``no talks'' & scripted flow \\
7 & viewer-count had zero callers (viewers poll; no socket to count) & cloud oracle \\
8 & four backend modules used a never-connected DB pool & HTTP probe \\
9 & \texttt{db.transaction()} called but unimplemented; non-atomic writes & exposed by \#8 \\
10 & welcome bonus granted twice once \#9 was fixed & exposed by \#9 \\
\bottomrule
\end{tabular}
\end{table}

% ==================================================================
\section{GUI Testing as a Kimiya Program}
\label{sec:as-kimiya}
% ==================================================================

We now make the correspondence exact. The claim is not analogy but
identity of structure: the harness is an informal, uncalibrated Kimiya
interpreter, and its scenarios are Kimiya programs whose datasheets were
left implicit.

\subsection{The mapping}
\label{sec:mapping}

\begin{table}[t]\small
\caption{Harness constructs are Kimiya constructs. The right column names
the datasheet quantity each carries.}
\label{tab:map}
\begin{tabular}{@{}lll@{}}
\toprule
Harness & Kimiya & Carries \\
\midrule
\textsf{locate}(desc, screen) & $\select\langle\rho_r\rangle(\text{desc},\text{screen})$ & recall $\rho_r$ \\
\textsf{assert\_screen}(desc) & $\judge\langle k,\tau\rangle\ \phi_\ctx(\text{screen})$ & $(\alpha,\beta,\rho)$ \\
\textsf{read\_text}(desc)$\to\$v$ & $\gen\langle\mathit{Str}\rangle(\text{screen})$, grounded downstream & validity $\nu$ \\
\texttt{verify\_db} / \texttt{verify\_cloud} & $\checkk\,\varphi$ (trusted kernel) & certainty $1$ \\
\texttt{verify\_slides\_match\_pdf} & $\checkk$ over a tolerance relation $\semeqk{\ctx_{\mathrm{img}}}$ & certainty $1$ \\
\texttt{verify\_audio\_out} & $\checkk\,\varphi$ (peak/silence predicate) & certainty $1$ \\
\texttt{click} retries, \texttt{optional:}, re-run-once & $\retry\ \cdot\ \mathbf{until}\ \cdot\ \mathbf{budget}\ \cdot$ & --- \\
scenario \textsc{fail} + artifacts & $\abstain$ / $\bolt$ (explicit, recorded) & --- \\
scenario \textsc{pass} & $\commitc$ of a triple $\{P\}\,\cdot\,\{Q\}_{\ge\conf}$ & --- \\
model calls per scenario & cost in the budget monoid & --- \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:map} is the core observation. Three entries deserve
comment.

\emph{Assertions are judgments under a purpose.} A harness assertion reads
``the live presentation view with a LIVE badge and an End Presentation
button.'' This is a rubric predicate $\phi_\ctx(\text{screen})$ whose
purpose $\ctx$ is implicit in the wording: it \emph{preserves} the named
widgets and key text, and \emph{allows loss} of theme, cursor, clock, and
unrelated windows. The harness's very robustness---the same assertion
passing in light and dark theme, at different window sizes, with the
operator's terminal visible beside the target---is exactly what a Kimiya
purpose licenses, and what a pixel-diff oracle forbids. Making $\ctx$
explicit (\S\ref{sec:transcribe}) is the discipline the harness was
missing.

\emph{Perceptual slide matching is a tolerance relation.} The oracle
\texttt{verify\_slides\_match\_pdf} accepts a stored slide $s$ as a
capture of source page $p$ when their perceptual hashes lie within a
distance threshold, $s\semeqk{\ctx_{\mathrm{img}}}p$. This relation is
reflexive, symmetric, and \emph{non-transitive}: two frames each within
distance~16 of a shared middle can be 24 apart. The oracle's design---
best-match against all pages plus a presentation-order constraint, rather
than positional $k\leftrightarrow k$ matching---is precisely a refusal to
chain the tolerance relation, the operational form of Kimiya's
non-transitivity axiom. We did not know we were obeying the axiom; the
axiom is what makes the oracle correct.

\emph{Reads are untrusted generation, trusted downstream.} When the
harness reads an 8-character group join code off browser~A's success
dialog into a variable and types it into browser~B, that read is a
$\gen$: untrusted. We never relied on it. Its correctness was established
\emph{downstream}, by B actually joining and a $\checkk$ against the
database membership count---exactly Kimiya's ``untrusted generator
proposes, trusted kernel disposes.''

\subsection{Why the harness almost never issued a false certificate}
\label{sec:false-cert}

The most consequential structural fact is why, across roughly thirty
runs, the vision layer produced \emph{zero} false passes. In Kimiya
terms, scenarios \emph{end in $\checkk$}. A judged guard
(\textsf{assert\_screen}, \textsf{locate}) gates \emph{progress}, not the
verdict. A judge false-accept ($\alpha$) merely lets the script proceed;
the committed claim---\texttt{cloud == kept}, per-talk sequence $1..N$, a
WAV peak above threshold---is $\checkk$-graded at certainty~1. So judge
error converts into wasted budget and abstention, almost never into a
false certificate. This is precisely what \textsc{Judge-If} composed with
\textsc{Check} predicts: the reliability that reaches the commit is the
product of a judge factor and a kernel factor, and when the kernel factor
is~1 on the load-bearing postcondition, the certificate is robust to
$\alpha$. It also explains the observed failure profile: failures were
overwhelmingly \emph{correct rejections}---a real bug, real state drift
from a prior failed run, or a wrong assertion---rather than silent
mispasses. The harness was, without knowing it, exploiting the
check/judge grade separation that is Kimiya's central safety property.

A corollary sharpens honesty. Not every harness postcondition is
kernel-graded. ``The ink stroke renders in browser~B'' and ``the LIVE
badge is shown'' are $\judge$-graded: certified only at $\beta$. A Kimiya
reading of a passing run therefore \emph{stratifies} its claims---which
hold at certainty~1 (counts, peaks, cloud totals) and which hold only at
the judge's operating point---where the current PASS conflates them. That
stratification is the paper's practical upgrade to the harness, and it is
where an undetected $\alpha$-error could still hide a product bug.

\subsection{A scenario, transcribed}
\label{sec:transcribe}

We transcribe the two-user collaboration scenario (create a group in A,
carry the join code to B, verify chat both ways) into Kimiya surface
syntax, purposes made explicit. $\actc$ and $\settle$
are the world-effecting constructs \emph{this paper proposes} in
\S\ref{sec:extend} (read $\actc$ as ``perform this input effect'');
they are typeset in \textcolor{extcolor}{\bfseries magenta} to
distinguish them from \textcolor{kwcolor}{\bfseries core Kimiya}
keywords, which appear in blue. Everything else is core Kimiya, with the
purpose of each $\select$ noted in a comment exactly as the Kimiya
overview does.

\begin{lstlisting}
context k_ui:                      -- locating a described control
  domain     = "GUI control matching a description on a screenshot"
  preserve   = [role, visible_label, affordance]
  allow_loss = [pixel_position, theme, scale, unrelated_windows]

context k_state:                   -- judging an application state
  domain     = "whether a screenshot shows a described UI state"
  preserve   = [key_widgets, key_text, key_status]
  allow_loss = [cursor, clock, theme, unrelated_windows]

-- A creates the group; the join code is generated (untrusted) and read.
act<A> click( select<0.97>("the + button by the groups heading", A.screen) )   -- under k_ui
act<A> click( select<0.97>("the Create Group button", A.screen) )               -- under k_ui
act<A> type ( "W3 Harness Group" )
act<A> click( select<0.97>("the Create submit button", A.screen) )              -- under k_ui
settle<A> until judge<1,1/1> shows(A.screen,
              "a success view with an 8-char join code") under k_state
          within 12s

code := gen<JoinCode>( read("the 8-character join code", A.screen) )   -- untrusted

-- B joins with the carried code.
act<B> click( select<0.97>("the + button by the groups heading", B.screen) )   -- under k_ui
act<B> click( select<0.97>("the Join Group button", B.screen) )                 -- under k_ui
act<B> click( select<0.97>("the Join Code input", B.screen) )                   -- under k_ui
act<B> type ( code );  act<B> key ( Return )
settle<B> until judge<1,1/1> shows(B.screen,
              "the group 'W3 Harness Group' is selected") under k_state
          within 12s

-- chat A -> B, then B -> A, each verified on the receiving screen ...
-- (omitted; identical shape)

-- THE VERDICT: kernel gates, then a judged gate, then commit.
if check( db.group_members("W3 Harness Group") = 2           -- kernel: certainty 1
        /\ db.messages(group) = [m_ab, m_ba] )               -- kernel: certainty 1
then if judge<1,1/1> shows(B.screen, "the chat pane displays m_ab")
        under k_state                                        -- judged: >= beta
     then commit(trace) else abstain
else abstain
\end{lstlisting}

Read the final block: it makes explicit what the informal PASS left
tacit. A run commits only through two gates. The kernel gate ($\checkk$
on the membership count and message rows) contributes no instrument
error: \emph{any} committing run satisfies those conjuncts with
certainty. The judged gate (``the chat pane displays $m_{ab}$'') is an
instrument reading at the locator's operating point; the certificate
carries that conjunct at the datasheet rate, not at $1$. A reviewer of the test---or an
LLM agent asked whether the collaboration feature is verified---learns
from the certificate exactly which guarantees are ironclad and which are
probabilistic, and at what rate.

\subsection{The reliability of a passing run, numerically}
\label{sec:numeric}

We instantiate the Kimiya logic on a passing scenario using the
\emph{measured} datasheet of \S\ref{sec:datasheet}. Consider a scenario
with $L$ load-bearing \textsf{locate} steps (each gating progress toward
the commit), $J$ judged guards on the path to the commit, and a
\texttt{commit} whose postcondition is a conjunction of kernel checks
$\varphi_{\mathrm{ker}}$ and judged claims $\psi_{\mathrm{jud}}$. Under the
harness retry policy (each locate retried, plus a scenario re-run on
terminal miss), the per-call locate recall $\rho_r$ composes to an
effective $\rho_r^{\mathrm{eff}}\approx 1$ over the campaign
(\S\ref{sec:datasheet}); we carry $\rho_r$ symbolically.

The kernel conjunct is certified at $1$ by \textsc{Check}. Each judged
guard on the path contributes its true-accept factor $\beta$ (the guard
must fire for the run to reach commit; \textsc{Judge-If} on the taken
branch). Each judged \emph{postcondition} claim contributes $\beta$
directly. So the certified reliability of the commit factors as
\[
\conf_{\mathrm{commit}}\ \ge\
\underbrace{\big(\rho_r^{\mathrm{eff}}\big)^{L}}_{\text{located}}\
\cdot\ \underbrace{\beta^{J}}_{\text{guards fired}}\
\cdot\ \underbrace{1}_{\varphi_{\mathrm{ker}}}\
\cdot\ \underbrace{\beta^{|\psi_{\mathrm{jud}}|}}_{\text{judged claims}} .
\]
For the transcribed scenario, the load-bearing kernel conjuncts (member
count, message rows) are the guarantee that matters, and they are
certified at $1$ regardless of $\beta$---this is the formal content of
\S\ref{sec:false-cert}. The \emph{judged} conjunct (``B's chat shows
$m_{ab}$'') is certified at $\beta\ge 0.976$ (\S\ref{sec:datasheet}). If
one instead demanded a single reliability figure for the \emph{whole}
commit read conjunctively, with $J=4$ guards on the path and one judged
postcondition, and taking the conservative $\beta=0.976$ and
$\rho_r^{\mathrm{eff}}=0.999$ over $L=8$ locates, the composed lower bound
is
\[
\conf_{\mathrm{commit}}\ \ge\ 0.999^{8}\cdot 0.976^{4}\cdot 1\cdot 0.976
\ =\ 0.992\cdot 0.908\cdot 0.976\ \approx\ 0.879,
\]
valid with datasheet-confidence $1-\delta$ (\S\ref{sec:datasheet}
publishes the locator sheet at the campaign's aggregate level). The
number is deliberately unglamorous, and that is the point: a green run is
not certainty, it is a priced guarantee, and the price is dominated by
how many \emph{judged} claims ride on the commit versus how many are
discharged to the kernel. The design lesson the arithmetic hands back to
the test author is exactly the one \S\ref{sec:false-cert} states: push
load-bearing postconditions into $\checkk$, and the $\beta^{(\cdot)}$
factors stop mattering.

% ==================================================================
\section{A Measured Datasheet for the Vision Locator}
\label{sec:datasheet}
% ==================================================================

A Kimiya instrument is only as trustworthy as its datasheet, and the
datasheet must be \emph{measured}, not assumed. We calibrate the harness's
vision locator (Claude Opus~4.8 via a headless CLI) from the campaign's
own retained artifacts: 58 scenario runs with structured step reports, and
the consoles of 55 runs recording per-call latency and attempt-level retry
lines. Ground truth is established in two tiers, reported separately:
\emph{(T1) outcome-grounding}---a step's latent truth inferred from the
run's terminal hard oracles, which are independent of the vision layer;
and \emph{(T2) per-event adjudication}---every non-passing perception
event in the campaign was root-caused when it occurred (product bug,
oracle bug, state drift, or model flake), and that record is reused here.
Where zero errors were observed we report the one-sided 95\% upper bound
(rule of three, $3/n$), never zero.

\begin{table}[t]\small
\caption{Measured operating point of the vision locator, from 58 runs /
604 model calls. Bounds are conservative (Wilson-95 lower for accept
rates; rule-of-three upper for unobserved error).}
\label{tab:sheet}
\begin{tabular}{@{}llll@{}}
\toprule
Instrument & Quantity & Value & Basis \\
\midrule
$\judge$ (assert) & $\beta$ true-accept & $\ge 0.976$ & $158/158$; Wilson-95 low \\
$\judge$ (assert) & $\alpha$ false-accept & $\le 0.16$ & $0/19$ true-neg; rule of three \\
$\select$ (locate) & $\rho_r$ per call & $\ge 0.975$ & $340/343$; Wilson-95 low \\
$\select$ (locate) & $\rho_r^{\mathrm{eff}}$ after retry & $\approx 1.000$ & $0$ unrecovered / 58 runs \\
$\gen$ (read) & value error & $\le 1.0$ & $0/3$; rule of three (thin) \\
cost & latency $p_{50}/p_{90}$ & $8.9/14.3$ s & 229 samples \\
cost & calls / scenario & $10.4$ & $604/58$ \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:sheet} is the result. Three points are worth stating
plainly, in the datasheet spirit of admitting what is weak.

\emph{The false-accept bound is the honest limitation.} We observed no
judge false-accept---including a blind re-read of six uniformly sampled
accepted assertions, all confirmed correct---but the campaign presented
only $\approx 19$ genuine true-negative trials to the assertion judge
(runs where a described state was truly absent and correctly rejected). The
rule of three therefore caps $\alpha\le 0.16$, no tighter. This is exactly
the quantity \S\ref{sec:false-cert} shows the harness is \emph{robust} to
(because commits end in $\checkk$), which is fortunate, because it is the
quantity we can least tightly bound from opportunistic data. A purpose-
built calibration suite---the standard Kimiya remedy---would drive it down.

\emph{Correlation is real and unmeasured.} The configuration is
single-model, single-prompt; immediate same-input retries are highly
correlated (we observed three consecutive misses on one plainly visible
element, then success on a later fresh run). We therefore treat an
immediate retry burst as \emph{one} effective sample and rely on the
scenario re-run---a genuinely fresh draw---for amplification. This is the
uncalibrated shadow of Kimiya's panel construct
(\lstinline|panel [A,B,C] paraphrase_prompts 2|): decorrelation by model
family and prompt paraphrase is what would let $k>1$ actually amplify, and
it is the first thing a production version should adopt.

\emph{Recall is the certificate-bearing direction.} Per-call locate recall
is high ($\ge 0.975$) and, under the retry policy, effectively $1$ over
the campaign (no unrecovered locate failure caused a false result). This
matches Kimiya's asymmetry: a locate \emph{miss} is loud (the step
retries or the scenario abstains, visibly), whereas a locate on the wrong
element is caught downstream when the subsequent judged/kernel step fails.
Recall, not precision, is what the certificate rides on, exactly as the
$\select$ datasheet demands.

% ==================================================================
\section{Extending Kimiya for Effectful, Timed Interaction}
\label{sec:extend}
% ==================================================================

The correspondence of \S\ref{sec:as-kimiya} is faithful for the
verification layer but strains at the interaction layer, because GUI
testing is Kimiya's deferred extension in its most demanding form:
computation that \emph{mutates an external world under adversarial
timing}. We identify three gaps and propose the constructs and proof
obligations a world-effecting Kimiya needs. We present this as design,
with the metatheory sketched, not mechanized.

\subsection{Gap 1: the store is an external, mutating world}

Kimiya's state is $\Sigma=\mathrm{Var}\rightharpoonup\mathbb{V}$: values
under the program's control. The harness's state is a live desktop, two
browsers, an audio server, and a production database---an external store
$\worldstore$ the program can \emph{act on} and \emph{observe} but not
directly read. We propose splitting the interaction primitives:
\begin{itemize}
\item $\actc\ e$: perform an input effect on $\worldstore$ (a click at a
located point, a keystroke, an audio injection). Its datasheet is a
\emph{delivery} contract---probability the effect is delivered, and its
cost---not a truth about the resulting state.
\item $\observe\langle S\rangle$: sample a schema-typed observation of
$\worldstore$ (a screenshot, a DB row, a waveform). An $\observe$ is the
\emph{subject} of a subsequent $\judge$ or $\checkk$; it is where latent
world-truth enters as an instrument reading.
\end{itemize}
The locate-then-click idiom becomes $\actc\
\mathsf{click}(\select\langle\rho_r\rangle(\text{desc},\observe(\text{screen})))$:
observe the screen, retrieve a point on it, act. Crucially, $\select$ now
retrieves over an \emph{observation}, so its recall datasheet is
conditioned on the observation family---which is exactly how the harness's
locator behaves (its reliability differs across the PyQt, GTK-dialog, and
browser surface families of Table~\ref{tab:sheet}).

\subsection{Gap 2: retry must reason about world frames}
\label{sec:worldretry}

Kimiya's \textsc{Retry} reruns the body \emph{from the loop's entry
state}---snapshot semantics---which is sound because the store is
program-owned and can be reset. An external world cannot be snapshotted: a
click that landed wrong may have opened a dialog or joined a call, and the
next attempt inherits it. This is not a corner case; it is the source of
the harness's hardest-won engineering (optional steps, leave-then-rejoin,
the ``dead microphone node'' hazard). A world-effecting retry needs a
\emph{frame condition} on the body:
\[
\ir{World-Retry}{
\begin{array}{c}
\{P\}\ C\ \{Q\}\ \text{on success} \qquad
C\ \text{is } \worldstore\text{-idempotent up to } \phi_{\mathrm{inv}} \\
(\alpha_{\mathrm{eff}},\beta_{\mathrm{eff}})\ \text{for } J \qquad
\text{on failure } C\ \text{runs } \mathsf{compensate}\ \text{restoring } \phi_{\mathrm{inv}}
\end{array}}
{\{P\wedge\phi_{\mathrm{inv}}\}\ \retry\ C\ \mathbf{until}\ J\
\mathbf{budget}\ \budget\ \{Q\}_{\ge(1-(1-g\beta_{\mathrm{eff}})^\budget)\pi_{\mathrm{acc}}}}
\]
Here $g$ is the body's datasheet task-family quality (per-round
probability of producing a $Q$-satisfying outcome) and
$\pi_{\mathrm{acc}}=g\beta_{\mathrm{eff}}/(g\beta_{\mathrm{eff}}+(1-g)\alpha_{\mathrm{eff}})$
the acceptance precision, both exactly as in Kimiya's \textsc{Retry}
rule~\citep{kimiya}; only the premises are new. They are the price of an
unresettable store. Either the body
is \emph{idempotent up to a declared world-invariant} $\phi_{\mathrm{inv}}$
(re-running a failed click leaves $\worldstore$ within $\phi_{\mathrm{inv}}$
of where it began---the property the harness's \texttt{optional:} steps in
informally assert), or it carries an explicit $\mathsf{compensate}$ action
that re-establishes $\phi_{\mathrm{inv}}$ before the next round (the
formal version of leave-then-rejoin). $\phi_{\mathrm{inv}}$ itself is a
$\checkk$- or $\judge$-checkable predicate on an $\observe$; when it is
judged, its error enters the reliability like any other. The obligation
that this premise cannot be met silently---that a body which mutates the
world outside $\phi_{\mathrm{inv}}$ is a \emph{type error}, not a runtime
surprise---is the interactive analogue of Kimiya's ban on silent semantic
equality.

\subsection{Gap 3: cost is timed, and time is adversarial}

Kimiya's cost monoid prices \emph{calls}; the harness's sharpest
methodology rule---never walk playback markers while audio plays, because
the locator's seconds of latency are wall-clock drift---has no home in it.
The world evolves \emph{while the judge thinks}. We propose a
$\settle\ \mathbf{until}\ J\ \mathbf{within}\ \delta$ construct: repeatedly
$\observe$-and-$\judge$ until $J$ holds or a real-time deadline $\delta$
elapses, with the deadline a first-class cost alongside the call budget.
$\settle$ is the principled form of the harness's \texttt{wait}$+$
\texttt{assert\_screen} idiom, and its obligation is a \emph{stability}
premise: the judged state must be one the world will hold still for at
least the judge's own latency, or the reading is of a state already gone.
The impossibility this formalizes---that no amount of judging can
certify a claim about a state that does not persist as long as the
judgment takes---is the timing analogue of the recall asymmetry: some
failures are silent unless the construct forbids them.

\subsection{What survives unchanged}
\label{sec:survives}

The verification layer is untouched by all three gaps. Purposes, the
check/judge grade split, budgets, abstention to $\bolt$, and audit
locality apply verbatim; the transcribed certificate of
\S\ref{sec:transcribe} is already written in the extended surface syntax
and its \texttt{commit} means exactly what the core logic says. The
extension is entirely in \emph{how state is touched}, not in \emph{how
reliability is proved}---which is why the harness could obey the
verification discipline (and reap its safety) while improvising the
interaction discipline (and paying for it in engineering).

\subsection{The extended calculus, assembled}
\label{sec:assembled}

Figure~\ref{fig:extgrammar} assembles the full language: core Kimiya
verbatim, plus the four world-effecting productions this paper proposes,
set in \textcolor{extcolor}{magenta} as everywhere in this paper. Three
disciplines are built into the grammar rather than left to convention.
First, \emph{the world is only ever touched through $\actc$ and read
through $\observe$}: no other production mentions $\worldstore$, so the
program-owned store and the external world cannot be silently conflated,
and $\select$ now ranges over an \emph{observation} (a value in the
store), never over the world directly---which is why its recall datasheet
is conditioned on the observation family (\S\ref{sec:datasheet}).
Second, every extended production carries its own contract: $\actc$ a
\emph{delivery} datasheet (probability the effect lands, and its cost)
that asserts nothing about the resulting state; $\observe$ a fidelity
and cost; $\settle$ a real-time deadline $\delta$ that is a first-class
cost alongside the call budget; and world-frame $\retry$ the invariant
and compensation obligations of \S\ref{sec:worldretry}. Third, a
\emph{freshness obligation} ties them together: the observation an
$\actc$ consumes (through the $\select$ that produced its target point)
must postdate the last effect on the same surface $\iota$---acting on a
stale screenshot is a type error, not a runtime surprise. This is the
grammatical form of the harness's practice of re-capturing the window
immediately before every click, and the timing analogue of Kimiya's ban
on silent semantic equality.

\begin{figure}[t]
\small
\begin{align*}
&\textit{surfaces} & \iota \Coloneqq{}& \mathsf{desktop} \mid A \mid B \mid \cdots
   && \text{declared regions of the world } \worldstore\\[2pt]
&\textit{effects} & \ext{a} \Coloneqq{}& \ext{\mathsf{click}(pt) \mid \mathsf{type}(s) \mid \mathsf{key}(k) \mid \mathsf{audio}(w) \mid \mathsf{net}(\mathit{up} \mid \mathit{down})}
   && \text{instance set, extensible}\\[2pt]
&\textit{commands} & C \Coloneqq{}& \Skip \;\mid\; x := e \;\mid\; x := \gen\langle S\rangle(p) \;\mid\; x := \select\langle\rho_r\rangle(q, x') \\
& & \mid{}& C_1 ; C_2 \;\mid\; \mathbf{if}\ J\ \mathbf{then}\ C_1\ \mathbf{else}\ C_2
      \;\mid\; \retry\ C\ \mathbf{until}\ J\ \mathbf{budget}\ \budget \\
& & \mid{}& \commitc(e) \;\mid\; \abstain \\
& & \mid{}& \ext{\actc^{\iota}\ a}
      && \ext{\text{delivery datasheet } (d_a, c_a)}\\
& & \mid{}& \ext{x := \observe^{\iota}\langle S\rangle}
      && \ext{\text{observation of } \worldstore \text{ at } \iota}\\
& & \mid{}& \ext{\settle^{\iota}\ \mathbf{until}\ J\ \mathbf{within}\ \delta}
      && \ext{\text{timed observe-and-judge}}\\
& & \mid{}& \ext{\retry\ C\ \mathbf{until}\ J\ \mathbf{budget}\ \budget\
      \{\mathbf{inv}\ \phi_{\mathrm{inv}};\ \mathbf{compensate}\ C_r\}}
      && \ext{\text{world-frame retry (\S\ref{sec:worldretry})}}\\[2pt]
&\textit{guards} & J \Coloneqq{}& \checkk\ \varphi \;\mid\; \judge\langle k,\tau\rangle\ \psi \\
&\textit{predicates} & \psi \Coloneqq{}& u \semeqk{\ctx} v \;\mid\; u \entailsk{\ctx} v \;\mid\; u \perp_{\ctx} v \;\mid\; \phi_{\ctx}(u)
\end{align*}
\caption{The world-effecting extension, assembled. Black productions are
core Kimiya, unchanged; \textcolor{extcolor}{magenta} productions are
this paper's extension. $\worldstore$ is the external world, touchable
only through $\actc$ and readable only through $\observe$; $\iota$
indexes declared surfaces (the harness's browser actors and desktop).
Judged predicates $\psi$ now take \emph{observations} as arguments---
latent world-truth enters the logic only as an instrument reading of an
$\observe$, at datasheet rates.}
\label{fig:extgrammar}
\end{figure}

% ==================================================================
\section{A Playbook for an Autonomous Testing Agent}
\label{sec:playbook}
% ==================================================================

The method is not tied to SeenSlide. We state it as a procedure an
autonomous language-model agent can follow to test arbitrary GUI software
under the same discipline. It is deliberately imperative and checkable;
each step names the Kimiya construct it produces.

\begin{enumerate}
\item \textbf{Fix the control flow first; never improvise it.} Write the
scenario as a deterministic step list (a user story: open $X$, do $Y$,
then $Z$). The agent may \emph{author} the script using its judgment, but
the \emph{run} must replay it exactly. A step list is a reviewable
artifact; an exploration trace is not.
\item \textbf{Make every meaning-laden step a named instrument.} Locating
a control is $\select$ (a description $+$ a recall expectation); judging a
state is $\judge$ (a description $+$ a purpose); reading a value is
$\gen$ (untrusted until grounded). Do not let the model choose the next
\emph{action}---only perceive.
\item \textbf{Declare a purpose for every judgment.} State, per assertion,
what must be preserved (the widgets/text that make the state that state)
and what may be lost (theme, position, clock, unrelated windows). This is
what makes the assertion robust and well-posed; an assertion without a
purpose is a pixel-diff in disguise.
\item \textbf{End every scenario in a non-visual kernel check.} A test
that concludes on a screen judgment certifies nothing at certainty~1.
Find the ground truth the feature is \emph{about}---a database row, a
public API response, a file on disk, an audio waveform, a log line---and
assert it with $\checkk$. This is the single most important rule: it is
what makes judge error convert to abstention rather than a false pass
(\S\ref{sec:false-cert}).
\item \textbf{Stratify the certificate.} In the \texttt{commit}, separate
kernel-graded conjuncts (certainty~1) from judged conjuncts
($\ge\beta$). Report both. A consumer of the test---human or agent---must
know which guarantees are ironclad.
\item \textbf{Retry with a declared world-invariant.} Wrap flaky
interaction in $\retry$/\texttt{optional}, but state what re-running is
allowed to disturb (\S\ref{sec:extend}): either the action is idempotent
up to $\phi_{\mathrm{inv}}$, or it carries a compensation. Treat a call
surviving from a prior run (a dialog, a joined call, a dead device
handle) as inherited world state, not a clean slate.
\item \textbf{Wait with $\settle$, not a fixed sleep, and never judge a
transient.} Poll-observe-judge until the state holds or a real-time
deadline passes. Do not perceive during continuous change (playback,
animation, streaming); the perception latency will read a state already
gone.
\item \textbf{Isolate and neutralize side effects.} Run the app under a
disposable environment (a fresh home directory, a throwaway identity, a
null credential store). For cloud tests, ensure the test identity cannot
touch real user data, and purge test artifacts with a filter that names
their test origin \emph{in the deletion itself}---never a list gathered
from logs. Route network through an injectable proxy so outages are a
test action, not a host change.
\item \textbf{Measure your instruments before you trust them.} Retain
every run's screenshots, judgments, and outcomes. Periodically
re-adjudicate a sample against the terminal kernel oracles to estimate
$(\alpha,\beta,\rho_r)$ and latency; publish the datasheet; recompute the
scenario's certified reliability (\S\ref{sec:numeric}). Report unobserved
error as a one-sided bound, never as zero.
\item \textbf{Decorrelate to amplify.} A single model retried on the same
prompt barely amplifies. When a judgment is load-bearing and cannot be
pushed to the kernel, vote across model families and prompt paraphrases
(a panel), and treat an immediate same-input retry burst as one effective
sample.
\end{enumerate}

Steps 4 and 5 are the discipline's core; a large-language-model agent that
follows only those two already tests far more safely than an autonomous
explorer, because its verdict stops depending on its own eyes. The
remaining steps buy tighter numbers and safer interaction.

% ==================================================================
\section{Toward Verified Interactive Systems}
\label{sec:verification}
% ==================================================================

The wider motivation is a route to \emph{formal verification of
interactive, networked, cloud-backed software}---systems that have long
resisted it for two reasons this decomposition addresses head-on.

Such systems resist verification because parts of their specification are
\emph{semantic}---``the viewer shows the presenter's current slide,''
``the audience hears the speaker,'' ``the schedule was imported
correctly''---and no decidable predicate captures them; and because their
state is an \emph{external world} no proof assistant can model in full. The
usual responses each fail one half: classical model checking demands the
whole system reduce to a finite transition system, which the semantics
and the world both refuse; and pure LLM-agent testing certifies nothing,
because an improvising agent's judgment is uncalibrated and its runs are
not artifacts.

A Kimiya program is a third option, and the harness is its existence
proof. It uses \emph{exactly as much determinism as the task requires}---
the control flow, the kernel oracles, the world-invariants---and
\emph{exactly as much language model as the task requires}---the
locates, the state judgments, the grounded reads---and the boundary
between them is not a matter of taste but of what each step \emph{can} be:
decidable predicates go to $\checkk$, meaning-laden ones to the
instruments. The certificate then states precisely what was verified at
certainty~1, what was verified at a measured error rate, and what was left
uncovered---and, by audit locality, its meaning depends only on the
program's declarations and the cited datasheets, so a reviewer (or a
downstream agent) can trust a test without replaying it.

Concretely, this suggests a research program. \emph{(i)} A
world-effecting Kimiya core (\S\ref{sec:extend}) with a mechanized
soundness theorem for \textsc{World-Retry} and $\settle$, reducing the
external world to an oracle-observed store with declared invariants.
\emph{(ii)} A calibration methodology for interface-perception instruments
---purpose-built suites per surface family (native toolkit, web, mobile),
published as datasheets an agent can cite. \emph{(iii)} A compiler from
user-story specifications to Kimiya test programs, closing the loop the
playbook opens by hand. The end state is a testing regime in which a
green run is not a boolean but a signed certificate---``this feature holds
at certainty~1 on these kernel facts and at $\ge\beta$ on these judged
facts, within this budget, citing these datasheets''---for exactly the
software that today has no verification story at all.

% ==================================================================
\section{Related Work}
\label{sec:related}
% ==================================================================

\emph{LLM-driven GUI testing} is predominantly mobile and autonomous:
scenario-guided and multi-agent exploration \emph{generate} actions,
where we execute fixed scripts and delegate only perception; desktop is a
recognized gap. \emph{Commercial AI test tools} offer natural-language
authoring and self-healing selectors, philosophically akin to our
$\select$, but scoped to web/mobile and silent on audio, native dialogs,
and cross-application flows. \emph{Multi-user feature testing} coordinates
LLM agents on separate mobile devices; we test two-user collaboration
deterministically, with the model reading a runtime-generated code across
browsers. \emph{Voice/WebRTC testing} evaluates STT/LLM/TTS pipelines,
typically via fake media streams; we found no system closing the audio
loop at the operating-system level inside a general GUI harness.
Foundationally, we build on Kimiya~\citep{kimiya}: our contribution is to
show that an independently constructed interactive test \emph{is} a Kimiya
program, to measure a real instrument datasheet, and to identify the
world-effecting extension the domain forces. (A full bibliography accompanies
the camera-ready; the companion technical report enumerates the harness's
ten bugs and scenarios in full.)

% ==================================================================
\section{Limitations}
\label{sec:limits}
% ==================================================================

The guarantees are conditional on the oracle model---judges being
calibratable instruments---which our datasheet supports in a restricted
regime (short, grounded, discriminative visual judgments) and no further.
The datasheet is measured on one product's interface families and one
model; the $\alpha$ bound is loose for want of true-negative trials
(\S\ref{sec:datasheet}), and correlation is treated conservatively rather
than measured. The world-effecting extension (\S\ref{sec:extend}) is a
design with sketched obligations, not a mechanized metatheory. The harness
is Linux/X11-specific in its input and capture layers, requires exclusive
use of the screen, and cannot test sub-second interactive behavior because
perception latency is seconds. Finally, our evidence is one system and one
operator: the ten bugs are an existence proof of effectiveness, not a
controlled comparison. What we claim is narrow and, we believe, solid:
that a real end-to-end interactive test, built without knowledge of the
theory, turned out to be a Kimiya program, and that reading it as one both
explains its safety and shows how to make it a certificate.

% ==================================================================
\section{Conclusion}
\label{sec:conclusion}
% ==================================================================

We set out to test a real interactive product and, in doing so, built a
harness that drives desktop and dual-browser surfaces, closed-loop audio,
network faults, and a live cloud from one deterministic step vocabulary,
finding ten shipped bugs. Reading the result through Kimiya, we found the
harness had implemented the theory unaware: its locator is $\select$, its
assertions are $\judge$ under implicit purposes, its reads are grounded
$\gen$, and its database and audio oracles are the $\checkk$ kernel whose
presence is precisely why judge error almost never became a false pass. We
measured the locator's operating point from the run artifacts, derived a
numeric certificate for a passing test, and identified the one place the
correspondence strains---the external, timed world---proposing the
constructs a world-effecting Kimiya needs. The method generalizes to a
playbook any language-model agent can follow. We read all of this as a
first, concrete step toward verifying the interactive, networked,
cloud-backed systems that have no verification story today: use exactly
the determinism and exactly the language model each step demands, and let
a Kimiya certificate make the division auditable.

\bibliographystyle{ACM-Reference-Format}
\begin{thebibliography}{1}
\bibitem[Anonymous(2026)]{kimiya}
Anonymous. 2026.
\newblock Kimiya: A Program Logic for Semantic Computation with Language Models.
\newblock \emph{Manuscript}.
\end{thebibliography}

\end{document}
